---
title: "Online Store Data Case Study"
author: "Hayden Lynch"
date: "2023-02-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fpp2)
library(forecast)
```

# Data Cleaning Process

### Step 1: Install/Import Libraries
```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(readr)
library(scales)
```

### Step 2: Import Unclean Data
```{r}
df = read_csv("online_store_customer_data_copy.csv")
df
```

### Step 3: Remove "NA" rows
```{r echo=FALSE, results='asis'}
library(knitr)
new_df = na.omit(df)
kable(df[1:10,], caption="A knitr kable")
```
* Using knitr kable's allow you to make nice tables in RMarkdown.

### Step 4: Remove all 2021 Rows
#### The "Transaction_date" needs to be in 'date format'.
```{r}
new_df = mutate(new_df, Transaction_date = as.Date(Transaction_date, "%m/%d/%Y"))
class(new_df$Transaction_date)
new_df
```
#### Now, rows including 2021 can be removed.
```{r}
new_df = new_df %>% 
  filter(Transaction_date < "2021-01-01")
new_df
```
* The tibble now has 1743 rows of data that can be analyzed.

### Step 5: Rename values in Referal column
```{r}
new_df$Referal[new_df$Referal == 1] = 'Reffered'
new_df$Referal[new_df$Referal == 0] = 'Not Referred'
new_df
```

### Step 6: Data Cleaning Process: Save cleaned data file
```{r}
save(new_df, file = "cleaned_data.csv")
write.csv(new_df, file = "cleaned_data.csv")
```

# Data Analysis

## Data Overview
```{r echo=TRUE, results='asis'}
df <- read_csv("cleaned_data.csv")
kable(df[1:10,], caption="Table Layout")
```
```{r}
colnames(df)
str(df)
glimpse(df)
```

## Data Analysis Process

### Question 1: Does the date influence spending amount (Year/Month)?
```{r echo=TRUE, results='asis'}
yearly_differences_amount_spent = df %>% 
  mutate(Year = format(Transaction_date, "%Y")) %>% 
  group_by(Year) %>% 
  summarise(Total_Amount_Spent = sum(Amount_spent))
kable(yearly_differences_amount_spent[1:2,])
```
* Per Year
```{r echo=TRUE, results='asis'}
monthly_differences_amount_spent = df %>% 
  mutate(Month = format(Transaction_date, "%m")) %>% 
  group_by(Month) %>% 
  summarise(Total_Amount_Spent = sum(Amount_spent))
kable(monthly_differences_amount_spent[1:12,])
```
* Per Month
```{r echo=TRUE, results='asis'}
date_differences_amount_spent = df %>% 
  mutate(Month = format(Transaction_date, "%m"), Year = format(Transaction_date, "%Y")) %>% 
  group_by(Month,Year) %>% 
  summarise(Total_Amount_Spent = sum(Amount_spent))
kable(date_differences_amount_spent[1:12,])
```
* Per both year and month

### Question 2: Do certain states spend more than others?
```{r echo=TRUE, results='asis'}
# Total amount spent per state
state_differences_amount_spent = df %>% 
  group_by(State_names) %>% 
  summarise(Total_Amount_Spent = sum(Amount_spent),
            Avg_Amount_Spent = mean(Amount_spent))
# Top performing states to the top of dataset
state_differences_amount_spent = state_differences_amount_spent[rev(order(state_differences_amount_spent$Total_Amount_Spent)),]
# Rounding to the second decimal
state_differences_amount_spent$Total_Amount_Spent = round(state_differences_amount_spent$Total_Amount_Spent,2)
state_differences_amount_spent$Avg_Amount_Spent = round(state_differences_amount_spent$Avg_Amount_Spent,2)
# View new dataset
kable(state_differences_amount_spent[1:5,])
```

### Question 3: Does marital status dictate membership segments?
```{r echo=TRUE, results='asis'}
# Total and average amount spent per marital status per membership
marital_status_dictate_membership = df %>% 
  group_by(Marital_status,Segment) %>% 
  summarise(Total_Amount_Spent = sum(Amount_spent),Avg_Amount_Spent = mean(Amount_spent))
marital_status_dictate_membership = marital_status_dictate_membership[order(marital_status_dictate_membership$Segment),]
kable(marital_status_dictate_membership[1:10,])
```

### Question 4: What is the percentage breakdown between employee status?
```{r echo=TRUE, results='asis'}
employee_status_percent = df %>% 
  group_by(Employees_status) %>% 
  count(Employees_status)
# Turning total count to percent of people (using scales package)
employee_status_percent$Percent = percent(employee_status_percent$n/1743)
#employee_status_percent$Percent = percent(employee_status_percent$Percent)
# Changing the name of column
names(employee_status_percent)[names(employee_status_percent) == 'n'] = "Count"
kable(employee_status_percent[1:4,])
```

### Question 5: What age group spends more than others, how does the payment method influence age group spending?
```{r echo=TRUE, results='asis'}
# Making Age_Group Column and new dataframe for the question
age_group_df = df %>% 
  mutate(
    Age_Group = dplyr::case_when(
      Age < 25 ~ "15-24",
      Age >= 25 & Age < 40 ~ "25-39",
      Age >= 40 & Age < 55 ~ "40-54",
      Age >= 55 ~ "55+"
      )
    )
age_group_differences_amount_spent = age_group_df %>% 
  group_by(Age_Group) %>% 
  summarise(Total_Amount_Spent = sum(Amount_spent), Avg_Amount_Spent = mean(Amount_spent))
# Get count of dataset
example = age_group_df %>% 
  count(Age_Group)
names(example)[names(example) == 'n'] = "Total People"
# Combine both tables
age_group_differences_amount_spent$Total_People = example$`Total People`
kable(age_group_differences_amount_spent[1:4,])
```

### Question 6: Are referrals worth investing into?
```{r echo=TRUE, results='asis'}
referal_amount_spent = df %>% 
  group_by(Referal) %>% 
  summarise(Total_Amount_Spent = sum(Amount_spent), Avg_Amount_Spent = mean(Amount_spent))
kable(referal_amount_spent[1:2,])
```

### Question 7: Should other payment methods be targeted and influenced?
```{r echo=TRUE, results='asis'}
payment_method_targeting = df %>% 
  group_by(Payment_method) %>% 
  summarise(Total_Amount_Spent = sum(Amount_spent), Avg_Amount_Spent = mean(Amount_spent))
kable(payment_method_targeting[1:3,])
```

### Question 8: How much of a difference are the different segments making?
```{r echo=TRUE, results='asis'}
# Creating table with total/avg amount spent per segment
segment_influence = df %>% 
  group_by(Segment) %>% 
  summarise(Total_Amount_Spent = sum(Amount_spent), Avg_Amount_Spent = mean(Amount_spent))
# Adding total people column
segment_count = df %>% 
  count(Segment)
segment_influence$Total_People = segment_count$n
# Adding percent column
segment_count$n = segment_count$n / 1743
segment_influence$Percent = segment_count$n
segment_influence$Percent = percent(segment_influence$Percent)
kable(segment_influence[1:5,])
```

### Question 9: In the varying states, which age group should be targeted, what percent do they make up in the states?
```{r echo=TRUE, results='asis'}
# Creating base table
state_targeting = age_group_df %>% 
  group_by(State_names, Age_Group) %>% 
  summarise(Total_Amount_Spent = sum(Amount_spent), Avg_Amount_Spent = mean(Amount_spent))
# Adding total people using count
state_agegroup_count = age_group_df %>% 
  group_by(State_names,Age_Group) %>% 
  count(Age_Group)
state_targeting$Total_People = state_agegroup_count$n
state_targeting = state_targeting[,c(1,2,5,3,4)]
# Sort by largest amount spent per state per age group
state_targeting = state_targeting %>% 
  arrange(desc(Total_Amount_Spent))
kable(state_targeting[1:10,])
```

### Question 10: Should we influence a gender for a specific segment?
```{r echo=TRUE, results='asis'}
# Creating table for total/avg amount spent
gender_segment_influence = df %>% 
  group_by(Gender,Segment) %>% 
  summarise(Total_Amount_Spent = sum(Amount_spent), Avg_Amount_Spent = mean(Amount_spent))
# Re-ordered the values based on segment
gender_segment_influence = gender_segment_influence[order(gender_segment_influence$Segment),]
# Created total people from different table using count
gender_segment_count = df %>% 
  group_by(Gender,Segment) %>% 
  count(Segment)
gender_segment_count = gender_segment_count[order(gender_segment_count$Segment),]
# After re-ordering the count, I merged both of them together
gender_segment_influence$Total_People = gender_segment_count$n
gender_segment_influence = gender_segment_influence[,c(1,2,5,3,4)]
kable(gender_segment_influence[1:10,])
```

### Question 11: What age group is worth referring to the online environment?
```{r echo=TRUE, results='asis'}
# Creating table for age_groups referred total/avg spending
age_group_online_experience = age_group_df %>% 
  group_by(Age_Group,Referal) %>% 
  summarise(Total_Amount_Spent = sum(Amount_spent), Avg_Amount_Spent = mean(Amount_spent))
# Separate table for total people
age_group_online_count = age_group_df %>% 
  group_by(Age_Group,Referal) %>% 
  count(Referal)
# Add column from count to main table
age_group_online_experience$Total_People = age_group_online_count$n
# Re-ordered table 
age_group_online_experience = age_group_online_experience[,c(1,2,5,3,4)]
kable(age_group_online_experience[1:8,])
```